---
title: "Untitled"
output:
  word_document: default
  pdf_document: default
  html_document:
    keep_md: yes
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

# Comparison between Exponential Distribution and Central Limit Theorem (CLT)

### David Pellon
### June 2017

## Overview:
### The objective of this study is comparing the exponential distribution and the Central Limit Theorem. We will use rexp(n, lambda) R function with lambda as the rate value = 0.2, a distribution of averages of 40 exponentials using a thousand simulations.

## Simulations:

### Definitions of the constants used: lambda, number of exponentials, number of simulations. Using a seed for reproducibility.

```{r a1, echo=TRUE}
lambda <- 0.2
seed   <- 12321
n_exp  <- 40
n_sim  <- 1000

set.seed(seed)
```

### Processing the 1000 averages of the 40 random exponentials

```{r a2, echo=TRUE}
means = NULL
for (i in 1: n_sim) means = c(means, mean(rexp(n_exp,lambda)))

```

## 1/ Comparing Sample Mean to the Theoretical Mean of the Distribution

### Sample mean:

```{r a3, echo=TRUE}
mean(means)
```
### Theoretical Mean of the distribution (lambda^-1):

```{r a4, echo=TRUE}
lambda^-1
```
### Comparison

```{r a5, echo=TRUE}
100*abs((lambda^-1) - mean(means))/mean(means)
```
The percentual difference is below 1%.

## 2/ Comparing Sample Variance and the Theoretical Variance of the Distribution

### Sample Variance:

```{r a6, echo=TRUE}
var(means)
```

### Theoretical Variance of the Distribution (lambda * sqrt(n_exp))^-2:

```{r a7, echo=TRUE}
(lambda * sqrt(n_exp))^-2
```

### Comparison:
```{r a8, echo=TRUE}
100*abs(var(means)-(lambda * sqrt(n_exp))^-2)/var(means)
```
The percentual difference is below 2%.

## 3/ Verifying if the difference between a large number of random exponentials and large number of averages of 40 exponentials is a normal distribution 

### Displaying a plot, comparing the theoretical mean of the distribution with the theoretical normal distribution for the 1000 simulations.

```{r a9, echo=TRUE}
library(ggplot2)
ggplot(data.frame(x=means),aes(x=means))+
  geom_histogram(binwidth=0.25,color='red',fill='white',aes(y=..density..)) +
  stat_function(fun=dnorm,args=list(mean=lambda^-1,sd=(lambda*sqrt(n_exp))^-1),size=2)+ labs(title="Verifying if the distribution is normal")+xlab("Sample Mean")+ylab("Density")
  
```











