---
title: "Untitled"
output:
  word_document: default
  pdf_document: default
  html_document:
    keep_md: yes
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

# Effect of Vitamin C on Tooth growth in Guinea Pigs

### David Pellon
### June 2017

## Overview:
### The objective of this study is comparing the tooth growth in Guinea Pigs depending on different diets and supplements.

## Load the ToothGrowth data and perform some basic exploratory data analyses:

### Loading data

```{r a1, echo=TRUE}
library(datasets)
data("ToothGrowth")
```

### Exploring data 

```{r a2, echo=TRUE}
head(ToothGrowth)
names(ToothGrowth)
summary(ToothGrowth)

```
We have 60 guinea pigs. Each group of them received different doses of Vitamin C using either orange juice or ascorbic acid.


```{r a3, echo=TRUE}
library(ggplot2)
levels(ToothGrowth$supp)<-c("Orange Juice","Ascorbic Acid")
ggplot(ToothGrowth,aes(x=factor(dose),y=len))+
  facet_grid(.~supp)+
  geom_boxplot(aes(fill=supp),show_guide=TRUE)+
  labs(title="Tooth Length, Dosage, Type of Supplement")+xlab("Dose (mg)")+ylab("Length")
  
```

Guinea Pigs which received the Vitamin using Orange Juice had greater length. The dose effect when using Ascorbic Acid had a more dramatic impact on tooth length than when using Orange Juice.


## Use confidence intervals/hypothesis to compare:

### With different doses, both supplements bring same tooth growth

For 0.5

```{r a4, echo=TRUE}
sset<-subset(ToothGrowth,dose ==0.5)
h05 <- t.test(len ~ supp, data=sset)
h05$conf.int
h05$p.value
```
Hypothesis with 0.5 mg, OJ Growth = AA Growth - FALSE
Hypothesis with 0.5 mg, OJ Growth > AA Growth - TRUE

For 1.0

```{r a5, echo=TRUE}
sset<-subset(ToothGrowth,dose ==1.0)
h10 <- t.test(len ~ supp, data=sset)
h10$conf.int
h10$p.value
```
Hypothesis with 1.0 mg, OJ Growth = AA Growth - FALSE
Hypothesis with 1.0 mg, OJ Growth > AA Growth - TRUE

For 2.0

```{r a6, echo=TRUE}
sset<-subset(ToothGrowth,dose ==2.0)
h20 <- t.test(len ~ supp, data=sset)
h20$conf.int
h20$p.value
```
Hypothesis with 2.0 mg, OJ Growth = AA Growth - NOT FALSE (Cannot be rejected)


## Conclusions

Orange Juice offers more tooth growth except for 2.0 dose that is not clear.
We assume that the data received is valid. 


